Introduction to *cellcount*

### 2022-10-05 EBImage_Markdown.Rmd
Initial upload
Updated

@@

*cellcount* is an R package distributed as part of the [Bioconductor](http://bioconductor.org) project. To install the package, start R and enter:

```{r, message=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
  install.packages("BiocManager")
}
if (!require("cellcount")) {
  BiocManager::install("cellcount")
}
library(cellcount)
```

The following packages are required. Please install and bring them into the R
environment: 
```{r, message=FALSE}
if (!require("EBImage")) {
  BiocManager::install("EBImage")
}
library(EBImage)
library(tiff)
library(pixmap)
library(raster)
library(beepr)
```

@@

You'll need to generate a couple of directories; one for saving the unique
color label images and another for CSV files. In this example, we are using 
images of *Microcystis* sp. cells. Please note that your 
directory addresses will be different. We recommend using a dedicated directory 
for *cellcount* results:

```{r,message=FALSE}
savdir <- ("C:/file path for saving CSV data")
image_savdir <- ("C:/file path for saving converted images")
```
```{r,include=FALSE}
savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_data/")
image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_results/")
```

@@

```{r,message=FALSE}
images <- list.files("C:/file path of images to analyze", pattern = "tif", full.name = T)
images_names <- list.files("C:/file path of images to analyze", pattern = "tif", full.name = F)
```
```{r,include=FALSE}
images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = T)
images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = F)
```

@@

```{r,include=FALSE}
greyscale <- function(x, contrast = 2) {
  x <- contrast * x
  x <- x[, , 1] + x[, , 2] + x[, , 3]
  x <- x / max(x)
  x <- normalize(x, inputRange = c(0.1, 0.75))
  return(x)
}
```
```{r}
grey_images <- lapply(img_transposed, greyscale, contrast = 1)
display(grey_images[[5]])
```

Next, we will use the "mapped" function to turn the 8-bit image we just
converted into a data matrix using the following code. This also includes a
pixel intensity threshold adjustment that end-users can change; this removes
any pixel that is under the threshold value to easily separate the background
from cells:

```{r,include=FALSE}
mapped <- function(x, threshold = 0.3) {
  x <- as.matrix(x)
  x[x < threshold] <- 0
  return(x)
}
```
```{r}
imagesMapped <- lapply(grey_images, mapped, threshold = 0.3)
```

Before we analyze our images in bulk, set aside some code space to visually 
analyze your images results. This could include adjusting the threshold
from the "mapped" function, the contrast from the "greyscale" function 
(if used), or could involve making adjustments to the "image_convert" function,
such as the width and height of the rectangular window, the pixel area
threshold (i.e., removes cells or other objects under a certain pixel area), 
or the tolerance and 'ext' variables. Observe your initial results as indicated
below and make adjustments where necessary:

```{r,include=FALSE}
border <- function(x, y) {
  c(x[1:y[1], 1], x[1:y[1], y[2]], x[1, 1:y[2]], x[y[1], 1:y[2]])
}
countCells <- function(x) {
  countAll <- max(x)
  dims <- dim(x)
  border1 <- border(x, dims)
  ids <- unique(border1[which(border1 != 0)])
  countInner <- countAll - length(ids)
  return(countInner)
}
countImages <- function(image, normalize = TRUE, removeEdgeCells = TRUE) {
  if (removeEdgeCells) {
    dims <- dim(image)
    border1 <- border(image, dims)
    ids <- unique(border1[which(border1 != 0)])
    inner <- rmObjects(image, ids)
    EBImage::colorLabels(inner, normalize = normalize)
  } else {
    EBImage::colorLabels(image, normalize = normalize)
  }
}
image_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
  image <- thresh(x, w = w, h = h, offset = offset)
  image1 <- fillHull(image)
  image2 <- watershed(distmap(image1), tolerance = tolerance, ext = ext)
  nf <- computeFeatures.shape(image2)
  nr <- which(nf[, "s.area"] < areathresh)
  image3 <- rmObjects(image2, nr)
  return(image3)
}
```
```{r}
imagesConverted <- image_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 0.8, ext = 1)
final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
display(final_img)
```

@@

```{r,include=FALSE}
outlier<-function(){
  mean<-median(Cell.Count$Cell_Count)
  sd<-sd(Cell.Count$Cell_Count)*2
  sd_neg<-(mean-sd)
  sd_pos<-(mean+sd)
  SD_range_detection<-ifelse(sd_pos>Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)&
    ifelse(sd_neg<Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)
  return(SD_range_detection)
}
```
```{r}
Cell.Count$Cell_Count<-as.numeric(Cell.Count$Cell_Count)
Cell.Count$SD_range<-outlier()
Cell.Count
```

@@

With the variables calculated, the following code uses these variables to 
calculate cell density and adds the data into the "Cell.Count" data frame:

```{r,include=FALSE}
cell_density <- function(x, FOV = 0.0726, images = 10, filtration.area = 213.8, volume = 1, total.volume = 5) {
  MCF <- filtration.area / FOV
  N <- x / images
  cyano <- (MCF * N) / volume
  cyano_total <- cyano * total.volume
  return(cyano_total)
}
```
```{r}
cell.den <- cell_density(cell.total, FOV = 0.0726, images = 10, filtration.area = 213.8, volume = 1, total.volume = 5)
Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Density", cell.den,"null")
```

@@

# Output data from test images in the *cellcount* package

```{r,echo=FALSE}
display(final_img)
Cell.Count
```
Test data results for *Microcystis* strain F-192

```{r,include=FALSE}
# clear environment and free unused memory
rm(list=ls())
gc()
```

```{r,include=FALSE}
greyscale <- function(x, contrast = 2) {
  x <- contrast * x
  x <- x[, , 1] + x[, , 2] + x[, , 3]
  x <- x / max(x)
  x <- normalize(x, inputRange = c(0.1, 0.75))
  return(x)
}
mapped <- function(x, threshold = 0.3) {
  x <- as.matrix(x)
  x[x < threshold] <- 0
  return(x)
}
border <- function(x, y) {
  c(x[1:y[1], 1], x[1:y[1], y[2]], x[1, 1:y[2]], x[y[1], 1:y[2]])
}
countCells <- function(x) {
  countAll <- max(x)
  dims <- dim(x)
  border1 <- border(x, dims)
  ids <- unique(border1[which(border1 != 0)])
  countInner <- countAll - length(ids)
  return(countInner)
}
countImages <- function(image, normalize = TRUE, removeEdgeCells = TRUE) {
  if (removeEdgeCells) {
    dims <- dim(image)
    border1 <- border(image, dims)
    ids <- unique(border1[which(border1 != 0)])
    inner <- rmObjects(image, ids)
    EBImage::colorLabels(inner, normalize = normalize)
  } else {
    EBImage::colorLabels(image, normalize = normalize)
  }
}
image_convert2 <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
  image <- thresh(x, w = w, h = h, offset = offset)
  image1 <- watershed(distmap(image), tolerance = tolerance, ext = ext)
  image2 <- fillHull(image1)
  nf <- computeFeatures.shape(image2)
  nr <- which(nf[, "s.area"] < areathresh)
  image3 <- rmObjects(image2, nr)
  return(image3)
}
cell_density <- function(x, FOV = 0.0726, images = 10, filtration.area = 213.8, volume = 1, total.volume = 5) {
  MCF <- filtration.area / FOV
  N <- x / images
  cyano <- (MCF * N) / volume
  cyano_total <- cyano * total.volume
  return(cyano_total)
}
```

```{r,echo=FALSE}
Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))

file_name<-paste0("img_")

savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_data/")
image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_results/")
images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/NZ_Anabena_60x/", pattern = "tif", full.name = T)
images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/NZ_Anabena_60x/", pattern = "tif", full.name = F)

imgNames <- paste0(file_name, images_names)
read_images <- lapply(images, readTIFF)
img_transposed <- lapply(read_images,aperm,c(2,1,3))
names(images) <- imgNames

grey_images <- lapply(img_transposed, greyscale, contrast = 1)

imagesMapped <- lapply(grey_images, mapped, threshold = 0.3)

imagesConverted <- image_convert2(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 0.8, ext = 1)
final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
display(final_img)
```

```{r,echo=FALSE}
for (z in 1:length(images)) {
  imagesConverted <- image_convert2(imagesMapped[[z]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 0.8, ext = 1)
  final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
  count <- countCells(imagesConverted)
  Cell.Count[nrow(Cell.Count) + 1, ] <- c(imgNames[[z]], count)
  analyzed_image <- paste0(sub(".tif", replacement = "", x = imgNames[z]), "_analyzed.tiff")
  writeImage(final_img, files = paste0(image_savdir, analyzed_image),compression=c("LZW"))
}

Cell.Count$Cell_Count<-as.numeric(Cell.Count$Cell_Count)
outlier<-function(){
  mean<-median(Cell.Count$Cell_Count)
  sd<-sd(Cell.Count$Cell_Count)*2
  sd_neg<-(mean-sd)
  sd_pos<-(mean+sd)
  SD_range_detection<-ifelse(sd_pos>Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)&
    ifelse(sd_neg<Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)
  return(SD_range_detection)
}
Cell.Count$SD_range<-outlier()

cell.total <- as.numeric(Cell.Count$Cell_Count)
cell.total <- sum(cell.total)

Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Total", cell.total,"null")

cell.den <- cell_density(cell.total, FOV = 0.027770534, images = 10, filtration.area = 213.8, volume = 0.5, total.volume =15)
Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Density", cell.den,"null")

cell.av<-(cell.total/10)
Cell.Count[nrow(Cell.Count) + 1, ] <- c("Imaging Average",cell.av,"null")

cell.mL<-(cell.den/15)
Cell.Count[nrow(Cell.Count) + 1, ] <- c("Total Volume Cell per mL",cell.mL,"null")

write.csv(Cell.Count, paste0(savdir, "/Anabena_test counts.csv"))

Cell.Count
```
Test data results for *Anabena* sp.

```{r,include=FALSE}
# clear environment and free unused memory
rm(list=ls())
gc()
```

```{r,include=FALSE}
greyscale <- function(x, contrast = 2) {
  x <- contrast * x
  x <- x[, , 1] + x[, , 2] + x[, , 3]
  x <- x / max(x)
  x <- normalize(x, inputRange = c(0.1, 0.75))
  return(x)
}
mapped <- function(x, threshold = 0.3) {
  x <- as.matrix(x)
  x[x < threshold] <- 0
  return(x)
}
border <- function(x, y) {
  c(x[1:y[1], 1], x[1:y[1], y[2]], x[1, 1:y[2]], x[y[1], 1:y[2]])
}
countCells <- function(x) {
  countAll <- max(x)
  dims <- dim(x)
  border1 <- border(x, dims)
  ids <- unique(border1[which(border1 != 0)])
  countInner <- countAll - length(ids)
  return(countInner)
}
countImages <- function(image, normalize = TRUE, removeEdgeCells = TRUE) {
  if (removeEdgeCells) {
    dims <- dim(image)
    border1 <- border(image, dims)
    ids <- unique(border1[which(border1 != 0)])
    inner <- rmObjects(image, ids)
    EBImage::colorLabels(inner, normalize = normalize)
  } else {
    EBImage::colorLabels(image, normalize = normalize)
  }
}
image_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
  image <- thresh(x, w = w, h = h, offset = offset)
  image1 <- fillHull(image)
  image2 <- watershed(distmap(image1), tolerance = tolerance, ext = ext)
  nf <- computeFeatures.shape(image2)
  nr <- which(nf[, "s.area"] < areathresh)
  image3 <- rmObjects(image2, nr)
  return(image3)
}
cell_density <- function(x, FOV = 0.0726, images = 10, filtration.area = 213.8, volume = 1, total.volume = 5) {
  MCF <- filtration.area / FOV
  N <- x / images
  cyano <- (MCF * N) / volume
  cyano_total <- cyano * total.volume
  return(cyano_total)
}
```

```{r,echo=FALSE}
Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))

file_name<-paste0("img_")

savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_data/")
image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_results/")
images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Shewanella_100x_DAPI/", pattern = "tif", full.name = T)
images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Shewanella_100x_DAPI/", pattern = "tif", full.name = F)

imgNames <- paste0(file_name, images_names)
read_images <- lapply(images, readTIFF)
img_transposed <- lapply(read_images,aperm,c(2,1,3))
names(images) <- imgNames

grey_images <- lapply(img_transposed, greyscale, contrast = 1)

imagesMapped <- lapply(grey_images, mapped, threshold = 0.3)

imagesConverted <- image_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
display(final_img)
```

```{r,echo=FALSE}
for (z in 1:length(images)) {
  imagesConverted <- image_convert(imagesMapped[[z]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
  final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
  count <- countCells(imagesConverted)
  Cell.Count[nrow(Cell.Count) + 1, ] <- c(imgNames[[z]], count)
  analyzed_image <- paste0(sub(".tif", replacement = "", x = imgNames[z]), "_analyzed.tiff")
  writeImage(final_img, files = paste0(image_savdir, analyzed_image),compression=c("LZW"))
}

Cell.Count$Cell_Count<-as.numeric(Cell.Count$Cell_Count)
outlier<-function(){
  mean<-median(Cell.Count$Cell_Count)
  sd<-sd(Cell.Count$Cell_Count)*2
  sd_neg<-(mean-sd)
  sd_pos<-(mean+sd)
  SD_range_detection<-ifelse(sd_pos>Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)&
    ifelse(sd_neg<Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)
  return(SD_range_detection)
}
Cell.Count$SD_range<-outlier()

cell.total <- as.numeric(Cell.Count$Cell_Count)
cell.total <- sum(cell.total)

Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Total", cell.total,"null")

cell.den <- cell_density(cell.total, FOV = 0.00042, images = 10, filtration.area = 213.8, volume = 0.02, total.volume =2)
Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Density", cell.den,"null")

cell.av<-(cell.total/10)
Cell.Count[nrow(Cell.Count) + 1, ] <- c("Imaging Average",cell.av,"null")

cell.mL<-(cell.den/2)
Cell.Count[nrow(Cell.Count) + 1, ] <- c("Total Volume Cell per mL",cell.mL,"null")

write.csv(Cell.Count, paste0(savdir, "/Shewanella_test counts.csv"))

Cell.Count
```
Test data results for *Shewanella* IRI-160

# Literature cited:

>>>>

# Overview:

Climate change and anthropogenic impacts have exacerbated the presence of 
harmful algal blooms (HABs) within marine and freshwater environments. 
Rising temperatures, stratification strength, reduction of vertical mixing, 
and increases in nutrients attribute to these increases in blooms. Routine 
monitoring of HABs has become important in order to understand ecosystem and 
public health risk and establish mitigation practices.

Standard methods involve the enumeration of HAB species using microscopy 
methods (i.e. hemocytometer), which is considered the gold standard, as well 
as other methods such as qPCR and flow cytometry. These methods can be time 
consuming and require significant knowledge related to these technologies. 
Technology has advanced enumeration needs to decrease time consumption while 
also maintaining accuracy. Instruments such as the FlowCam&trade; and Imaging 
Flow Cytobot (IFCB) have been developed for these purposes; however, these 
instruments are expensive to purchase, operate, and maintain. The need for 
open-source technologies is important for accessibility reasons, but also for 
end-user customization needs.

Our project sought to solve the two mentioned issues on developing enumeration 
technology for open-source capabilities: accessibility and customization. A 
prior report released by Pokrzywinski et al. (2019) had outlined the 
development of an automated cyanobacteria enumeration script using the 
open-source programming language, R, along with several functions from the 
R package *EBImage*. However, the previous presentation of 
this script can be considered extremely difficult to grasp for inexperienced 
R users. We utilized this script as a template for the development of an R 
package, involving the reduction of code into simple functions, expanding 
the ability for end-user modifications, development of a vignette for 
inexperienced users, and promoting open-source availability. Along with these 
goals, we aim to establish a rapid and accurate enumeration method for 
researchers, where significantly less time is consumed compared to the 
standard enumeration methods available.

@@

The following packages are required. Please install and bring them into the R
environment: 
```{r, message=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
  install.packages("BiocManager")
}
if (!require("EBImage")) {
  BiocManager::install("EBImage")
}
library(EBImage)
library(tiff)
library(pixmap)
library(raster)
```

@@

You'll need to generate a couple of directories; one for saving the unique
color label images and another for CSV files. In this example, we are using 
images of *Microcystis* sp. cells that are contained within the package as test
data. Please note that your directory addresses will be different. We recommend 
using a dedicated directory for *cellcount* results:

```{r,message=FALSE}
savdir <- ("file path for saving CSV data")
image_savdir <- ("file path for saving converted images")
```
```{r,include=FALSE}
savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_data/")
image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_results/")
```

@@

```{r,message=FALSE}
images <- list.files("file path of images to analyze", pattern = "tif", full.name = T)
images_names <- list.files("file path of images to analyze", pattern = "tif", full.name = F)
```
```{r,include=FALSE}
images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = T)
images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = F)
```

@@

```{r}
greyscale <- function(x, contrast = 2) {
  x <- contrast * x
  x <- x[, , 1] + x[, , 2] + x[, , 3]
  x <- x / max(x)
  x <- normalize(x, inputRange = c(0.1, 0.75))
  return(x)
}
```
```{r}
grey_images <- lapply(img_transposed, greyscale, contrast = 1)
display(grey_images[[5]])
```

Next, we will use the "mapped" function to turn the 8-bit image we just
converted into a data matrix (using 'as.matrix') using the following code. 
This also includes a pixel intensity threshold adjustment that end-users can 
change; this removes any pixel that is under the threshold value to easily 
separate the background from cells:

```{r}
mapped <- function(x, threshold = 0.3) {
  x <- as.matrix(x)
  x[x < threshold] <- 0
  return(x)
}
```
```{r}
imagesMapped <- lapply(grey_images, mapped, threshold = 0.3)
```

Before we analyze our images in bulk, we introduce here several functions
that are required to take matrix data to count cells within images and to
produce 'color-label' images. These include the following:

**"countImages" function:** Displays converted microscopy images of distinct 
cell counts subsequently from the image_convert function. Can use the "display" 
function from EBImage to automatically generate the produced image from this 
function.

```{r}
countImages <- function(image, normalize = TRUE, removeEdgeCells = TRUE) {
  if (removeEdgeCells) {
    dims <- dim(image)
    border1 <- border(image, dims)
    ids <- unique(border1[which(border1 != 0)])
    inner <- rmObjects(image, ids)
    EBImage::colorLabels(inner, normalize = normalize)
  } else {
    EBImage::colorLabels(image, normalize = normalize)
  }
}
```

**"border" function:** Takes images to predict cells along a predetermined 
border in order to count cells within the predetermined border. Cells along the
predetermined border are not counted, which is outlined in the "countCells"
function.

```{r}
border <- function(x, y) {
  c(x[1:y[1], 1], x[1:y[1], y[2]], x[1, 1:y[2]], x[y[1], 1:y[2]])
}
```

**"countCells" function:** Takes information from a converted image and counts 
cells within the image. This uses the "border" function to remove cells along 
the image border and only count cells within the image.

```{r}
countCells <- function(x) {
  countAll <- max(x)
  dims <- dim(x)
  border1 <- border(x, dims)
  ids <- unique(border1[which(border1 != 0)])
  countInner <- countAll - length(ids)
  return(countInner)
}
```

**"image_convert" function:** Takes a mapped image (i.e., data matrix) and 
counts individual cells within an image. This is to be used with a single image 
with cells that contains a single fluorescent dimension 
(i.e. auto-fluorescence, DAPI, etc.). This is useful for analyzing images
containing single-cell cyanobacteria.

```{r}
image_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
  image <- thresh(x, w = w, h = h, offset = offset)
  image1 <- fillHull(image)
  image2 <- watershed(distmap(image1), tolerance = tolerance, ext = ext)
  nf <- computeFeatures.shape(image2)
  nr <- which(nf[, "s.area"] < areathresh)
  image3 <- rmObjects(image2, nr)
  return(image3)
}
```

We also include a separate, yet similar function to "image_convert" called 
"image_convert2". This function is ideally meant for filamentous or 
colonial-type cyanobacteria that are harder to discern with the standard 
"image_convert" function. The difference between these functions are that 
the "watershed" function from the EBImage package comes first in the outlined 
function compared to the "image_convert" function.

```{r}
image_convert2 <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
  image <- thresh(x, w = w, h = h, offset = offset)
  image1 <- watershed(distmap(image), tolerance = tolerance, ext = ext)
  image2 <- fillHull(image1)
  nf <- computeFeatures.shape(image2)
  nr <- which(nf[, "s.area"] < areathresh)
  image3 <- rmObjects(image2, nr)
  return(image3)
}
```

In addition, set aside some code space to visually 
analyze your images results. This could include adjusting the threshold
from the "mapped" function, the contrast from the "greyscale" function 
(if used), or could involve making adjustments to the "image_convert" function,
such as the width and height of the rectangular window, the pixel area
threshold (i.e., removes cells or other objects under a certain pixel area), 
or the tolerance and 'ext' variables. Observe your initial results as indicated
below and make adjustments where necessary:

```{r}
imagesConverted <- image_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 250, 
                                 tolerance = 0.8, ext = 1)
final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
display(final_img)
```

@@

```{r}
outlier<-function(){
  mean<-median(Cell.Count$Cell_Count)
  sd<-sd(Cell.Count$Cell_Count)*2
  sd_neg<-(mean-sd)
  sd_pos<-(mean+sd)
  SD_range_detection<-ifelse(sd_pos>Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)&
    ifelse(sd_neg<Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)
  return(SD_range_detection)
}
```
```{r}
Cell.Count$Cell_Count<-as.numeric(Cell.Count$Cell_Count)
Cell.Count$SD_range<-outlier()
Cell.Count
```

@@

With the variables calculated, the following "cell_density" uses these 
variables to calculate cell density and adds the data into the "Cell.Count" 
data frame:

```{r}
cell_density <- function(x, FOV = 0.0726, images = 10, filtration.area = 213.8, 
                         volume = 1, total.volume = 5) {
  MCF <- filtration.area / FOV
  N <- x / images
  cyano <- (MCF * N) / volume
  cyano_total <- cyano * total.volume
  return(cyano_total)
}
```
```{r}
cell.den <- cell_density(cell.total, FOV = 0.0726, images = 10, filtration.area = 213.8, 
                         volume = 1, total.volume = 5)
Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Density", cell.den,"null")
```

@@

#Literature cited:
